On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-15 TTau_Halpha/kurosawa.tex
Change made in Discussion sections. The rearranges made in the order of presentation in Model section.  Plus many minor changes.  (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@617 25464f21-6099-402c-b30a-640f1d40d38b

We present  radiative transfer models of classical T~Tauri stars, and
investigate the formation of H$\alpha$ line emission 
in order to understand their complex circumstellar environments, and to explain
the wide variety of line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with magnetospherical accretion alone to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind) presented,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al, and discuss the basic physical conditions required to reproduce
the profiles in each classified type.

@@

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt*{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt*{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet*{takami:2003}
show direct evidence for the presence of bipolar and monopolar
outflows down to $\sim 1\,\mathrm{AU}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observations using high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc both above and below the disc, suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observations
of HH30 \citep{burrows:1996} show the jet traced to within $\lesssim30$~AU
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 AU \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their sample (30 CTTS) show blue-shifted
absorption components in at least one of the Balmer lines and Ca~K (most
commonly in H$\alpha$).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model space is divided
into four different regions: (1)~a central continuum source, (2)~the magnetospherical
accretion flow, (3)~the wind outflow, and (4)~the accretion
disc. Fig.~\ref{fig:config_dipolar} 
depicts the relative location of the regions in the model space. In
all regions, the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas falls freely, moving along the magnetic field onto the
surface of the star. For simplicity,
the collimated bipolar wind is defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described. In our models, two different
types of the wind configurations are considered: (1)~the bipolar
wind resembling the density structure of the MHD simulation by \citet{krasnopolsky:2003}
in large scale ($\sim 100$~AU) in section~\ref{sub:wind_jet}, and
(2) that in small scale ($\sim 10$~AU) in section~\ref{sub:Alternative-model-disc-wind}.

@@

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{AU}\right]$\tabularnewline
\hline
Standard&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

@@

As mentioned earlier, the wind model presented here is to simulate
the collimated MHD disc-wind model of \citet{krasnopolsky:2003} on a 
large scale ($\sim 100$~AU) , and an alternative wind model which
resembles the model of \citet{krasnopolsky:2003} in small scale ($\sim
10$~AU)
will be presented later (section~\ref{sub:Alternative-model-disc-wind}).

@@

\subsection{Alternative model: the disc-wind model}

@@

The wind model presented here is intended to mimic the collimated MHD disc-wind
calculations of \citet{krasnopolsky:2003} in small scale ($\sim 10$~AU).
\citet*{knigge:1995} introduced the {}``split-monopole'' kinematic
disc-wind model in their studies of the UV resonance lines formed
in the winds of cataclysmic variable stars. In this model, the outflow
arises from the surface of the rotating accretion disc, and has a
biconical geometry. The specific angular momentum is assumed to be
conserved along a stream line, and the poloidal velocity component
is assumed to be simply a radial from vertically displaced {}``sources''
from the central star. We adopt their disc-wind model here, and consider
the combined effect of the disc-wind and the magnetospherical accretion
flow model described earlier. In the following, we briefly described
their disc-wind model. For details of the models, readers are referred
to \citet{knigge:1995} and \citet{long:2002}.

Four basic parameters for the models are: (1) mass-loss rate, (2)~the
degree of the wind collimation, (3)~the velocity gradients, and (4)~the
wind temperature. The basic configuration of the disc-wind model is
shown in Figure~\ref{fig:config_discwind}. The disc wind originates
from the disc surface, but the {}``source'' point ($S$), from which
the stream lines diverges, are placed at distance $d$ above and below
the centre of the star. By changing the value of $d$, the degree
of the wind collimation is controlled. The larger the value of $d$,
the wind becomes more parallel to the rotation axis, i.e. the wind
becomes more collimated. The mass-loss launching is restricted between
$R_{\mathrm{wi}}$ and $R_{\mathrm{wo}}$ where the former is set
to be equal to the outer radius of the magnetosphere ($R_{\mathrm{mo}}$)
for simplicity and the latter is set to $1$~AU as in \citealt{krasnopolsky:2003}.

@@

\caption{Dependency of the disc-wind density and velocity on the wind
acceleration parameter $\beta$. The wind density $\rho$ (left panel)
and the poloidal velocity component $V_{p}$ (right panel) along the
stream line starting from the mid point of the wind launching zone,
i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
\right ) / 2$, are shown as a function of the distance ($l$) from the
wind launching point (c.f.~
equations~\ref{eq:discwind-poloidal-velocity} and
\ref{eq:discwind-density}). The azimuthal velocity component
($V_{\theta}$), which is independent of $\beta$
(c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
the right panel for a comparison. The density is normalised to the
density $\rho_{0}$ at the wind launching point for $\beta = 1.0$ case.
The $v_{p}$ reaches the terminal velocity by 100~AU for all
$\beta$. In the far field ($l>10$~AU), the density is approximately
proportional to $\sim l^2$. Up to $l~\sim 10$~AU, the density is
smaller and the poloidal speed is larger for the wind with a larger
$\beta$ value. }

@@

\caption{Dependency of H$\alpha$ profiles on inclination ($i$).  The
profiles are computed with the magneospherical accretion flow using
the standard parameters given in Table~\ref{tab:std_parameters} and
$T_{\mathrm{max}}=7500~\Kelvin$. The solid, dotted, and dashed lines
are for $i=10^{\circ}$, $55^{\circ}$, and $80^{\circ}$
respectively. As the inclination becomes larger, the peak flux and the
equivalent width of the line becomes smaller.  Similar dependency is
seen in the models with the wide ranges of $T_{\mathrm{max}}$ and
$\beta$. }

@@

\caption{H$\alpha$ model profiles computed with wide ranges of the
  isothermal wind temperature ($T_{\mathrm{wind}}$) and the wind
  accerelation rate ($\beta$). The profiles are computed with the wind
  component only (no magnetospherical accretion flow or accretion
  disc). The `standard' model parameters (Table~\ref{tab:std_parameters}) are
  applied. The wind mass-loss rate ($\dot{M}_{\mathrm{wind}}$),
  the inclination ($i$) and the degree of collimination ($b$) are
  fixed at $10^{-8}~{\mathrm{M_{\sun}~yr^{-1}}}$, $55^{\circ}$ and
  $4.0$\, respectively. The temperature (indicated along the vertical
  axis) of the model increases from top to bottom, and the wind
  accerelation rate increases from left to right. The P~Cygni profile
  are seen for lower $T_{\mathrm{wind}}$ and lower $\beta$ (slow
  accerelation) models.  The postion of the absorption the component
  moves toward the line centre as $\beta$ increases. The effect of
  line broadening becomes more prominant for higher $T_\mathrm{wind}$
  and higher $\beta$ models.}

@@

\includegraphics[%
  clip,
  scale=0.45]{figures/windonly_B_effect2.eps}

@@

\caption{The effect of wind collimation. The line profiles computed
with with $\left(T_{\mathrm{wind}},~\beta \right) = \left(
9000\Kelvin, 0.5\right)$ for the wind anisotropy parameter $b=0, 2, 4$
are shown in dashed, dotted, and solid respectively.  All the other
parameters are same as in Figure~\ref{fig:atlas_wind_model}.  The
P~Cygni profile feature becomes prominant for the spherical wind model
since the column density of along the line of a sight of the observer
at a mid inclination angle ($i=55^{\circ}$) increases for lower $b$
values for a fixed mass-loss rate of the wind. The blue-ward asymmetry
is still present even in the isotropic wind case ($b=0$).}

@@

The bipolar nature of the wind is clearly seen in the images. In this
example, the most of the wind emission occurs within a few to a several
(depending mainly on the value of $\beta$) stellar radii from the
star. The figure also shows that the accretion disc blocking the wind
emission from the bottom half of the wind which is moving away from
the observer. This results in the reduction of the wind emission in
the red wing. The image (mid panel) also shows that the accretion
disc is not occulting the stellar disc at this inclination; therefore,
the ratio of the wind emission flux to the photospherical continuum
flux becomes slightly smaller for the model with the disc. This results
in the reduction in the peak flux and the flux in the wings in the
normalised profile shape. If the disc becomes more flared (if the
opening angle exceeds $45^{\circ}$at the outer edge of the accretion
disc), the occultation of the stellar disc by the accretion could
be seen at this inclination angle. Although it is not shown here,
the continuum level of the two models are identical. For the models
with $i<30^{\circ}$, the presence of the disc has little/no effect
on the profile shapes. In the figures, the extent of the H$\alpha$
emission is smaller ($\sim 0.2\,\mathrm{AU}$ or $\sim 20\, R_{*}$)
compared to the spectro-astrometric observations of \citet{takami:2003}
for RU~Lup and CS Cha which show 1--5~AU scale outflows%
\footnote{\citet{takami:2003} their observation also show that some objects
(e.g~Z CMa and AS~353A) displaying the outflows of larger scale
($>50\,\mathrm{AU}$); however, this could be formed by shocks rather
than MHD-wave heating (e.g. \citealt{hartmann:1982}) or X-ray heating
(e.g.~\citealt{shang:2002}). %
}; however, the extent of the line emission region becomes slightly
larger ($\sim0.4$~AU) if a slower wind acceleration rate ( $\beta=4$)
is used.

Although not shown here, we have also computed the H$\alpha$ profiles
from the combination of the collimated wind and the accretion disc
($R_{\mathrm{di}}=2.2R_{*}$ and $R_{\mathrm{do}}=100\,\mathrm{AU}$) for 
the same range of $\beta$ and $T_{\mathrm{max}}$ values used in 
Figure~\ref{fig:atlas_wind_model}.  As expected
(c.f.~Figure~\ref{fig:effect_alpha_disc}), the models show the 
decrease in the flux in the blue wing when compared to the flux of the
same model computed without the accretion disc.
The overall behaviour of the line profile shapes as functions of $\beta$ and $T_{\mathrm{wind}}$
are very similar to that for the models without the accretion disc.
A significant reduction of the flux in the red wing is seen in the
wind model with the accretion disc when compared with the wind only
model. The wings appear to be weaker in these models compared to the
wind only models because the ratio of the wind emission flux
to the continuum flux is lower.

@@

\caption{A comparison of the wind models with and without the
optically-thick and geometrically-thin disc.  The latter can be also
equivalent to a model with a large inner radius of the accretion disc
in which none of the wind emission region is not blocked provided the
inclination is moderate. The left panel shows the profile computed
with $i=55^{\circ}$, $\beta=0.5$ and $T_{\mathrm{wind}} =
9000~\mathrm{K}$ (see text for other parameters used). The flux in the
red wing of the line profile from the model with the disc (solid) is
greatly reduced by the presence of the accretion disc compared to that
from the model without the disc (dashed).  The peak flux and the flux
in the wings are also reduced in the model with the disc because the
disc does block the wind emission region, but none of the stellar
disc, i.e. the continuumm flux level is unchanged.  The H$\alpha$
image (project in the plane of the observer) for the model with the
disc (middle panel) and that for the model without (right plane) are
also shown. The image flux is computed by integrating the specific
flux over the wavelength range ($-800\,\kmps < v < 800\,\kmps$. The
flux is in an arbitrary units and on a logarithmic (base $10$) scale. }

@@

\caption{The effect of the mass-loss rate to mass-accretion rate ratio
($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The H$\alpha$
profile computed with only the magnetopsherical accretion (thick
solid) --- $T_{\mathrm{max}}=7500\,\Kelvin$,
$\dot{M}_{acc}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$ --- is compared with
the magnetosphere, wind, and disc hybrid models for $\mu =$ 0.01
(solid), 0.025 (dash-dot), 0.1 (dot) and 0.2 (dash). All the models
are computed with the inclination angle $i=55^{\circ}$. As $\mu$
increases the P-Cygni absorption deepens and the emission component
becomes stronger. In general, the absorption is too strong for the
models with $\mu>0.025$ compared to the observations (e.g.,
\citealt{reipurth:1996}).}

@@

\caption{Same as Figure~\ref{fig:atlas_wind_model}, but the models
include the magnetospherical accretion flow (with
$T_{\mathrm{\max}}=7500~\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$),  and the
accretion disc. The mass-loss rate of the wind is
$\dot{M}_{\mathrm{wind}}=10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$. A
relatively small mass-loss rate is chosen to avoid unrealistically
strong P-cyg absorption component (see text in
section~\ref{sub:model-acc-disc-wind} and
Figure~\ref{fig:effect_mdot_disc_wind_acc}).  For a comparison, the
profile computed with only the magensopherical accretion flow (dotted)
is overlayed. All the profiles are computed at the inclination angle
$i=55^{\circ}$. As $\beta$ becomes larger, the postion of the wind
absorption component moves toward the line centre. Although little
wind emission is seen in this temperature range, it becomes more
important in a model with a higher temperature and a larger wind
mass-loss rate. The profiles resembles the observed H$\alpha$ emission
profiles affected by the wind.}

@@

Although the profiles shown in Figure~\ref{fig:atlas_disc_wind_acc_model}
are those at a single inclination angle ($i=55^{\circ}$), they are
very similar to the Type II-B and III-B of the observed CTTS H$\alpha$
profiles (\citealt{reipurth:1996}), which cannot be explained
by the magnetospherical accretion flow models (e.g.~\citealt{muzerolle:2001}).
Comparison of the model profiles with different types of the observed
H$\alpha$ profiles will be given in section~\ref{sub:Classification-scheme-proposed}.

@@

\caption{Sample H$\alpha$ model profiles which charactersize the
classfication by \citet{reipurth:1996}. The combination of
magnetospherical accretion flow, the accretion disc, and the disc wind
model can reasonably reproduce wide ranges of H$\alpha$ profiles seen
in observations. Similar results can be obtained if the bipolar wind is
replaced by the disc wind. The model parameters and short comments are
summarised in Table~\ref{tab:classification}. The horizontal axes
are velocities in $\kmps$, and the vertical axes are nomalized
flux.}

>>>>

We present  radiative transfer models of classical T~Tauri stars, and
investigate the formation of H$\alpha$ line emission 
in order to understand their complex circumstellar environments, and to explain
the wide variety of line profiles seen in observations. To
overcome the limited applicability of the radiative transfer models
with magnetospherical accretion alone to fitting observed H$\alpha$
profiles, and to any other lines affected by the stellar wind,
two types of kinematic wind models are introduced: (1)~the bipolar
wind model in which the wind is originated from the star itself, and
(2)~the disc-wind model in which the wind originates from the inner
part of the accretion disc. We perform systematic investigations of
the model parameters for the wind and the magnetosphere to search
for possible geometrical and physical conditions which lead to the
types of profiles seen in observations. We find both wind models
can reproduce the wide range profile types seen is observations, 
although the inclination dependence of the line equivalent width
predicted by the bipolar wind model agree with trends seen in the
observation, but the disc-wind model does not. Using the model results,
we examine the H$\alpha$ spectroscopic classification used by Reipurth
et.~al, and discuss the basic physical conditions required to reproduce
the profiles in each classified type.

@@

Many observational studies (e.g., \citealt{herbig:1962}; \citealt{edwards:1994};
\citealt{kenyon:1994}; \citealt*{reipurth:1996}; \citealt{alencar:2000})
of CTTS line profiles show evidence for both outward wind flows and
inward accretion flows, as seen in the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC) profiles
respectively. Typical mass-loss rates of CTTS are about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$
to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ (e.g., \citealt{kuhi:1964};
\citealt{edwards:1987}; \citealt*{hartigan:1995}), and the mass-accretion
rates are also about $10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$
(e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988}; \citealt{gullbring:1998}).
Recent H$\alpha$ spectro-astrometric observations by \citet*{takami:2003}
show direct evidence for the presence of bipolar and monopolar
outflows down to $\sim 1\,\mathrm{au}$ scale (e.g.\,CS~Cha and RU~Lup).
Similarly, ESO VLT observations using high-resolution ($R=50\,000$)
two-dimensional spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B,
and HV~Tau~C) by \citet{appenzeller:2005} show the extended H$\alpha$
emission in the direction perpendicular to the obscuring circumstellar
disc both above and below the disc, suggesting the presence
of the bipolar outflows. On a larger scale, \emph{HST} observations
of HH30 \citep{burrows:1996} show the jet traced to within $\lesssim30$~au
from the star. The jet has a cone shape with an opening angle of $3^{\circ}$
between 70 and 700 au \citep{koenigl:2000}. \citet{alencar:2000}
found about 80 per cent of their sample (30 CTTS) show blue-shifted
absorption components in at least one of the Balmer lines and Ca~K (most
commonly in H$\alpha$).

@@

To understand how the different part of the CTTS circumstellar environment
contributes to the formation of H$\alpha$, the model space is divided
into four different regions: (1)~a central continuum source, (2)~the magnetospherical
accretion flow, (3)~the wind outflow, and (4)~the accretion
disc. Fig.~\ref{fig:config_dipolar} 
depicts the relative location of the regions in the model space. In
all regions, the density is symmetric around the z-axis. The innermost radius of
the magnetosphere at the equatorial plane coincides with the the inner
radius of the accretion disc. From the inner most part of the accretion
disc, the gas falls freely, moving along the magnetic field onto the
surface of the star. For simplicity,
the out-flowing wind component is defined only in the outside of the
largest radius of the magnetosphere (to avoid overlapping of the wind
with the magnetosphere). In the following subsections, the details
of model components will be described.
\begin{figure*}

@@

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
Standard&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

@@

\subsection{Outflow geometries}
Two different types of the wind configurations are considered: (1)~the
bipolar wind model which resembles the density structure of the MHD
simulation by \citet{krasnopolsky:2003} in \em large scale \em ($\sim
100$~au), and (2)~the disc-wind model which resembles the density of
the same MHD simulation in \em small scale \em ($\sim 10$~au).  While
the wind density is more concentrated toward the polar directions in
(1), it is more concentrated toward the equatorial direction in
(2). The following describes each model in details.

\subsubsection{Bipolar Wind}

@@

@@

\subsubsection{Disc wind}

@@

The wind model presented here is intended to mimic the collimated MHD disc-wind
calculations of \citet{krasnopolsky:2003} in small scale ($\sim 10$~au).
\citet*{knigge:1995} introduced the {}``split-monopole'' kinematic
disc-wind model in their studies of the UV resonance lines formed
in the winds of cataclysmic variable stars. In this model, the outflow
arises from the surface of the rotating accretion disc, and has a
biconical geometry. The specific angular momentum is assumed to be
conserved along a stream line, and the poloidal velocity component
is assumed to be simply a radial from vertically displaced {}``sources''
from the central star. We adopt their disc-wind model here, and consider
the combined effect of the disc-wind and the magnetospherical accretion
flow model described earlier. In the following, we briefly described
their disc-wind model. For details of the models, readers are referred
to \citet{knigge:1995} and \citet{long:2002}.

Four basic parameters for the models are: (1) mass-loss rate, (2)~the
degree of the wind collimation, (3)~the velocity gradients, and (4)~the
wind temperature. The basic configuration of the disc-wind model is
shown in Figure~\ref{fig:config_discwind}. The disc wind originates
from the disc surface, but the {}``source'' point ($S$), from which
the stream lines diverges, are placed at distance $d$ above and below
the centre of the star. By changing the value of $d$, the degree
of the wind collimation is controlled. The larger the value of $d$,
the wind becomes more parallel to the rotation axis, i.e. the wind
becomes more collimated. The mass-loss launching is restricted between
$R_{\mathrm{wi}}$ and $R_{\mathrm{wo}}$ where the former is set
to be equal to the outer radius of the magnetosphere ($R_{\mathrm{mo}}$)
for simplicity and the latter is set to $1$~au as in \citealt{krasnopolsky:2003}.

@@

\caption{Dependency of the disc-wind density and velocity on the wind
acceleration parameter $\beta$. The wind density $\rho$ (left panel)
and the poloidal velocity component $V_{p}$ (right panel) along the
stream line starting from the mid point of the wind launching zone,
i.e. $\left ( w,z \right ) = \left ( w_{\mathrm{mid}}, 0 \right )$
where $w_{\mathrm{mid}} = \left( R_{\mathrm{wi}}+R_{\mathrm{wo}}
\right ) / 2$, are shown as a function of the distance ($l$) from the
wind launching point (c.f.~
equations~\ref{eq:discwind-poloidal-velocity} and
\ref{eq:discwind-density}). The azimuthal velocity component
($V_{\theta}$), which is independent of $\beta$
(c.f. equation~\ref{eq:discwind-toroidal-velocity}), is also shown in
the right panel for a comparison. The density is normalised to the
density $\rho_{0}$ at the wind launching point for $\beta = 1.0$ case.
The $v_{p}$ reaches the terminal velocity by 100~au for all
$\beta$. In the far field ($l>10$~au), the density is approximately
proportional to $\sim l^2$. Up to $l~\sim 10$~au, the density is
smaller and the poloidal speed is larger for the wind with a larger
$\beta$ value. }

@@

\caption{Dependency of H$\alpha$ profiles on inclination ($i$).  The
profiles are computed with the magnetospherical accretion flow using
the standard parameters given in Table~\ref{tab:std_parameters} and
$T_{\mathrm{max}}=7500~\Kelvin$. The solid, dotted, and dashed lines
are for $i=10^{\circ}$, $55^{\circ}$, and $80^{\circ}$
respectively. As the inclination becomes larger, the peak flux and the
equivalent width of the line becomes smaller.  Similar dependency is
seen in the models with the wide ranges of $T_{\mathrm{max}}$ and
$\beta$. }

@@

\caption{H$\alpha$ model profiles computed with wide ranges of the
  isothermal wind temperature ($T_{\mathrm{wind}}$) and the wind
  acceleration rate ($\beta$). The profiles are computed with the wind
  component only (no magnetospherical accretion flow or accretion
  disc). The `standard' model parameters (Table~\ref{tab:std_parameters}) are
  applied. The wind mass-loss rate ($\dot{M}_{\mathrm{wind}}$),
  the inclination ($i$) and the degree of columniation ($b$) are
  fixed at $10^{-8}~{\mathrm{M_{\sun}~yr^{-1}}}$, $55^{\circ}$ and
  $4.0$\, respectively. The temperature (indicated along the vertical
  axis) of the model increases from top to bottom, and the wind
  acceleration rate increases from left to right. The P~Cygni profile
  are seen for lower $T_{\mathrm{wind}}$ and lower $\beta$ (slow
  acceleration) models.  The position of the absorption the component
  moves toward the line centre as $\beta$ increases. The effect of
  line broadening becomes more prominent for higher $T_\mathrm{wind}$
  and higher $\beta$ models.}

@@

\includegraphics[%
  clip,
  scale=0.62]{figures/windonly_B_effect2.eps}

@@

\caption{The effect of wind collimation. The line profiles computed
with with $\left(T_{\mathrm{wind}},~\beta \right) = \left(
9000\Kelvin, 0.5\right)$ for the wind anisotropy parameter $b=0, 2, 4$
are shown in dashed, dotted, and solid respectively.  All the other
parameters are same as in Figure~\ref{fig:atlas_wind_model}.  The
P~Cygni profile feature becomes prominent for the spherical wind model
since the column density of along the line of a sight of the observer
at a mid inclination angle ($i=55^{\circ}$) increases for lower $b$
values for a fixed mass-loss rate of the wind. The blue-ward asymmetry
is still present even in the isotropic wind case ($b=0$).}

@@

The bipolar nature of the wind is clearly seen in the images. In this
example, the most of the wind emission occurs within a few to a several
(depending mainly on the value of $\beta$) stellar radii from the
star. The figure also shows that the accretion disc blocking the wind
emission from the bottom half of the wind which is moving away from
the observer. This results in the reduction of the wind emission in
the red wing. The image (mid panel) also shows that the accretion
disc is not occulting the stellar disc at this inclination; therefore,
the ratio of the wind emission flux to the photospherical continuum
flux becomes slightly smaller for the model with the disc. This results
in the reduction in the peak flux and the flux in the wings in the
normalised profile shape. If the disc becomes more flared (if the
opening angle exceeds $45^{\circ}$at the outer edge of the accretion
disc), the occultation of the stellar disc by the accretion could
be seen at this inclination angle. Although it is not shown here,
the continuum level of the two models are identical. For the models
with $i<30^{\circ}$, the presence of the disc has little/no effect
on the profile shapes. In the figures, the extent of the H$\alpha$
emission is smaller ($\sim 0.2\,\mathrm{au}$ or $\sim 20\, R_{*}$)
compared to the spectro-astrometric observations of \citet{takami:2003}
for RU~Lup and CS Cha which show 1--5~au scale outflows%
\footnote{\citet{takami:2003} their observation also show that some objects
(e.g~Z CMa and AS~353A) displaying the outflows of larger scale
($>50\,\mathrm{au}$); however, this could be formed by shocks rather
than MHD-wave heating (e.g. \citealt{hartmann:1982}) or X-ray heating
(e.g.~\citealt{shang:2002}). %
}; however, the extent of the line emission region becomes slightly
larger ($\sim0.4$~au) if a slower wind acceleration rate ( $\beta=4$)
is used.

Although not shown here, we have also computed the H$\alpha$ profiles
from the combination of the collimated wind and the accretion disc
($R_{\mathrm{di}}=2.2R_{*}$ and $R_{\mathrm{do}}=100\,\mathrm{au}$) for 
the same range of $\beta$ and $T_{\mathrm{max}}$ values used in 
Figure~\ref{fig:atlas_wind_model}.  As expected
(c.f.~Figure~\ref{fig:effect_alpha_disc}), the models show the 
decrease in the flux in the blue wing when compared to the flux of the
same model computed without the accretion disc.
The overall behaviour of the line profile shapes as functions of $\beta$ and $T_{\mathrm{wind}}$
are very similar to that for the models without the accretion disc.
A significant reduction of the flux in the red wing is seen in the
wind model with the accretion disc when compared with the wind only
model. The wings appear to be weaker in these models compared to the
wind only models because the ratio of the wind emission flux
to the continuum flux is lower.

@@

\caption{A comparison of the wind models with and without the
optically-thick and geometrically-thin disc.  The latter can be also
equivalent to a model with a large inner radius of the accretion disc
in which none of the wind emission region is not blocked provided the
inclination is moderate. The left panel shows the profile computed
with $i=55^{\circ}$, $\beta=0.5$ and $T_{\mathrm{wind}} =
9000~\mathrm{K}$ (see text for other parameters used). The flux in the
red wing of the line profile from the model with the disc (solid) is
greatly reduced by the presence of the accretion disc compared to that
from the model without the disc (dashed).  The peak flux and the flux
in the wings are also reduced in the model with the disc because the
disc does block the wind emission region, but none of the stellar
disc, i.e. the continuum flux level is unchanged.  The H$\alpha$
image (project in the plane of the observer) for the model with the
disc (middle panel) and that for the model without (right plane) are
also shown. The image flux is computed by integrating the specific
flux over the wavelength range ($-800\,\kmps < v < 800\,\kmps$. The
flux is in an arbitrary units and on a logarithmic (base $10$) scale. }

@@

\caption{The effect of the mass-loss rate to mass-accretion rate ratio
($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The H$\alpha$
profile computed with only the magnetospherical accretion (thick
solid) --- $T_{\mathrm{max}}=7500\,\Kelvin$,
$\dot{M}_{acc}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$ --- is compared with
the magnetosphere, wind, and disc hybrid models for $\mu =$ 0.01
(solid), 0.025 (dash-dot), 0.1 (dot) and 0.2 (dash). All the models
are computed with the inclination angle $i=55^{\circ}$. As $\mu$
increases the P-Cygni absorption deepens and the emission component
becomes stronger. In general, the absorption is too strong for the
models with $\mu>0.025$ compared to the observations (e.g.,
\citealt{reipurth:1996}).}

@@

\caption{Same as Figure~\ref{fig:atlas_wind_model}, but the models
include the magnetospherical accretion flow (with
$T_{\mathrm{\max}}=7500~\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$),  and the
accretion disc. The mass-loss rate of the wind is
$\dot{M}_{\mathrm{wind}}=10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$. A
relatively small mass-loss rate is chosen to avoid unrealistically
strong P-Cyg absorption component (see text in
section~\ref{sub:model-acc-disc-wind} and
Figure~\ref{fig:effect_mdot_disc_wind_acc}).  For a comparison, the
profile computed with only the magnetospherical accretion flow (dotted)
is overlayed. All the profiles are computed at the inclination angle
$i=55^{\circ}$. As $\beta$ becomes larger, the position of the wind
absorption component moves toward the line centre. Although little
wind emission is seen in this temperature range, it becomes more
important in a model with a higher temperature and a larger wind
mass-loss rate. The profiles resembles the observed H$\alpha$ emission
profiles affected by the wind.}

@@

Although the profiles shown in
Figure~\ref{fig:atlas_disc_wind_acc_model} are those at a single
inclination angle ($i=55^{\circ}$), they are very similar to the most
common types of H$\alpha$ profiles seen in the observations
(e.g.~\citealt{reipurth:1996}). Apparently, this type of profile
(affected by the blue-shifted absorption feature) cannot be explained
by the magnetospherical accretion flow alone
(e.g.~\citealt{muzerolle:2001}).  Comparison of the model profiles
with different types of the observed H$\alpha$ profiles will be given
in section~\ref{sub:Classification-scheme-proposed}.

@@

\caption{Sample H$\alpha$ model profiles which characterise the
classification by \citet{reipurth:1996}. The combination of
magnetospherical accretion flow, the accretion disc, and the disc wind
model can reasonably reproduce wide ranges of H$\alpha$ profiles seen
in observations. Similar results can be obtained if the bipolar wind is
replaced by the disc wind. The model parameters and short comments are
summarised in Table~\ref{tab:classification}. The horizontal axes
are velocities in $\kmps$, and the vertical axes are normalised
flux.}
